Early Observations of the Type Ia Supernova \abc:\\ A Case of Interaction with Nearby, Unbound Material and/or Strong Ejecta Mixing

### 2018-05-22 doc/ms.tex
Add discussion section

The proliferation of high-cadence, time-domain surveys has led to several SNe
Ia being discovered within $\sim$2\,d of first light in the past $\sim$decade.
Observations probing the early evolution of these SNe allows us to place
unique constraints on their progenitor systems and the corresponding explosion
physics. This has revealed diversity in the earliest epochs after explosion,
and that commonly used SNe Ia templates do not match observations at these
phases (e.g., \citealt{2012ApJ...744...38F}).

In \S\ref{sec:first_light} we compared our early observations of \abc\ to
SN\,2011fe, which has the most comprehensive observations of the young SN Ia
sample. Given its normal spectroscopic and photometric evolution, SN\,2011fe
has been adopted as a standard for the early evolution of SNe Ia in many
studies. While a detailed quantitative analysis is beyond the scope of this
study, a qualitative examination of very young SNe Ia that otherwise exhibit
normal spectra and evolution at and post peak\footnote{This definition
excludes iPTF\,14atg, which was shown to be subluminous with SN\,2002es-like
spectra \citep{2015Natur.521..328C}.} reveals considerable diversity. In other
words, at early times SN\,2011fe may not be the norm.

For SN\,2011fe the initial rise is well described by a $t^2$ power law, the
$(B - V)_0$ colors evolve from the red to the blue in the $\sim$week after
explosion, and the \ion{C}{2} present in the initial spectra is weak
\citep{2011Natur.480..344N,2016ApJ...820...67Z,2012ApJ...752L..26P}. In
contrast, \abc\ exhibits a near linear rise in flux, the $(B - V)_0$ colors
are blue and roughly constant, and the \ion{C}{2} absorption is strong.
Examining just these 3 qualitative features, SN\,2009ig is well matched to
SN\,2011fe \citep{2012ApJ...744...38F}, while SN\,2013dy
\citep{2013ApJ...778L..15Z}, SN\,2017cbv \citep{2017ApJ...845L..11H}, and
\abc\ all bear a striking resemblance. SN\,2012cg, on the other hand, is
intermediate to these two groups with weak C II and a relatively shallow early
rise, like SN\,2011fe, but blue $(B - V)_0$ colors, like \abc\
\citep{2012ApJ...756L...7S,2016ApJ...820...92M}. SN\,2014J is intermediate in
the other direction in that it exhibits a near linear rise
\citep{2014ApJ...783L..24Z,2015ApJ...799..106G}, but the color evolution is
very similar to SN\,2011fe \citep{2014ApJ...788L..21A}.\footnote{\ion{C}{2} is
not detected in the spectra of SN\,2014J
\citep{2014ApJ...784L..12G,2014ApJ...783L..24Z}, though the earliest spectra
of SN\,2014J were obtained at a much later phase than the other SNe discussed
here. \citet{2015ApJ...798...39M} find evidence for \ion{C}{1} in the NIR
spectra of SN\,2014J, but this detection cannot constrain the relative
strength of \ion{C}{2} and \ion{Si}{2} shortly after explosion.} That these
early observations cannot be easily separated into two distinct groups
suggests that it is unlikely that a single physical mechanism drives the
diversity of SNe Ia at early times.

In the case of SN\,2012cg and SN\,2017cbv it has been argued that the early
blue optical colors are indicative of interaction between the SN ejecta and a
binary companion \citep{2016ApJ...820...92M,2017ApJ...845L..11H}. SN\,2017cbv
is particularly remarkable in that the observations presented in
\citet{2017ApJ...845L..11H} show a clearly resolved bump in the $U$, $B$, and
$g'$ bands in the $\sim$5\,d after explosion. In
\citeauthor{2017ApJ...845L..11H} it is found that the bump can be explained
via the combination of ejecta--companion interaction and the normal evolution
of a SN Ia. A challenge for this model, similar to \abc\ (see
\S\ref{sec:companion}), is that it significantly overpredicts the UV
brightness of the SN compared to what is observed. Indeed, in the case of
SNe\,2012cg, 2017cbv, and \abc\ the $\mathrm{UV} - \mathrm{optical}$ colors
are significantly redder than those observed in iPTF\,14atg. It is argued in
\citet{2017ApJ...845L..11H} that several model assumptions, including: (i)
ideal blackbody emission, (ii) a constant opacity, (iii) a simple power-law
density profile for the ejecta, and (iv) spherical symmetry, may be incorrect,
which could reconcile the discrepancy with the UV observations. Above, we
argued for interaction with diffuse, unbound material and strong
$^{56}$Ni mixing as a possible explanation for \abc, and indeed
\citet{2017ApJ...845L..11H} consider these possibilities for SN\,2017cbv as
well. Separately, several arguments against companion interaction for
SN\,2012cg are presented in \citet{2016arXiv161007601S}.

Ultimately, there are arguments in favor of and against each of the
possibilities to model the early emission from SNe Ia. Moving
forward, more detailed models and simulations are needed to properly explain
the observed diversity. No matter the correct explanation for the early
behavior of SNe\,2013dy, 2017cbv, and \abc, the strong similarities between
these events suggests that they may reflect a common physical origin.

>>>>

The proliferation of high-cadence, time-domain surveys has led to several
SNe Ia being discovered within $\sim$2\,days of first light in roughly the
past decade. Observations probing the early evolution of these SNe allow us
to place unique constraints on their progenitor systems and the
corresponding explosion physics. This has revealed diversity in the earliest
epochs after explosion, and that commonly used SN Ia templates do not match
observations at these phases (e.g., \citealt{2012ApJ...744...38F}).

In \S\ref{sec:first_light} we compared our early observations of \abc\ to
SN\,2011fe, which has the most comprehensive observations of the young SN Ia
sample. Given its normal spectroscopic and photometric evolution, SN\,2011fe
has been adopted as a standard for the early evolution of SNe Ia in many
studies. While a detailed quantitative analysis is beyond the scope of this
study, a qualitative examination of very young SNe Ia that otherwise exhibit
normal spectra and evolution at peak and post-peak\footnote{This definition
excludes iPTF\,14atg, which was shown to be subluminous with SN\,2002es-like
spectra \citep{2015Natur.521..328C}.} reveals considerable diversity. In
other words, at early times SN\,2011fe may not be the norm.

For SN\,2011fe the initial rise is well described by a $t^2$ power law, the
$(B - V)_0$ colors evolve from the red to the blue in the $\sim$1\,week
after explosion, and the \ion{C}{2} present in the initial spectra is weak
\citep{2011Natur.480..344N,2012ApJ...752L..26P,2016ApJ...820...67Z}. In
contrast, \abc\ exhibits a near-linear rise in flux, the $(B - V)_0$ colors
are blue and roughly constant, and the \ion{C}{2} absorption is strong.
Examining just these three qualitative features, SN\,2009ig is well matched
to SN\,2011fe \citep{2012ApJ...744...38F}, while SN\,2013dy
\citep{2013ApJ...778L..15Z}, SN\,2017cbv \citep{2017ApJ...845L..11H}, and
\abc\ all bear a striking resemblance. SN\,2012cg, on the other hand, is
intermediate to these two groups, with weak \ion{C}{2} and a relatively
shallow early rise, like SN\,2011fe, but blue $(B - V)_0$ colors, like \abc\
\citep{2012ApJ...756L...7S,2016ApJ...820...92M}. SN\,2014J is intermediate
in the other direction in that it exhibits a near-linear rise
\citep{2014ApJ...783L..24Z,2015ApJ...799..106G}, but the color evolution is
very similar to that of SN\,2011fe
\citep{2014ApJ...788L..21A}.\footnote{\ion{C}{2} is not detected in the
spectra of SN\,2014J \citep{2014ApJ...784L..12G,2014ApJ...783L..24Z}, though
the earliest spectra of SN\,2014J were obtained at a much later phase than
the other SNe discussed here. \citet{2015ApJ...798...39M} find evidence for
\ion{C}{1} in the NIR spectra of SN\,2014J, but this detection cannot
constrain the relative strength of \ion{C}{2} and \ion{Si}{2} shortly after
explosion.} That these early observations cannot be easily separated into
two distinct groups suggests that it is unlikely that a single physical
mechanism drives the diversity of SNe Ia at early times.

In the case of SN\,2012cg and SN\,2017cbv it has been argued that the early
blue optical colors are indicative of interaction between the SN ejecta and
a binary companion \citep{2016ApJ...820...92M,2017ApJ...845L..11H}.
SN\,2017cbv is particularly remarkable in that the observations presented in
\citet{2017ApJ...845L..11H} show a clearly resolved bump in the $U$, $B$,
and $g'$ bands in the $\sim$5\,days after explosion. In
\citeauthor{2017ApJ...845L..11H} it is found that the bump can be explained
via the combination of ejecta--companion interaction and the normal
evolution of a SN Ia. A challenge for this model, similar to \abc\ (see
\S\ref{sec:companion}), is that it significantly overpredicts the UV
brightness of the SN compared to what is observed. Indeed, in the case of
SN\,2012cg, SN\,2017cbv, and \abc\ the $\mathrm{UV} - \mathrm{optical}$
colors are significantly redder than those observed in iPTF\,14atg. It is
argued in \citet{2017ApJ...845L..11H} that several model assumptions,
including (i) ideal blackbody emission, (ii) a constant opacity, (iii) a
simple power-law density profile for the ejecta, and (iv) spherical
symmetry, may be incorrect, which could reconcile the discrepancy with the
UV observations. Above, we argued for interaction with diffuse, unbound
material and strong $^{56}$Ni mixing as a possible explanation for \abc, and
indeed \citet{2017ApJ...845L..11H} consider these possibilities for
SN\,2017cbv as well. Separately, several arguments against companion
interaction for SN\,2012cg are presented in \citet{2016arXiv161007601S}.

Ultimately, there are arguments in favor of and against each of the
possibilities to model the early emission from SNe Ia. Moving forward, more
detailed models and simulations are needed to properly explain the observed
diversity. No matter the correct explanation for the early behavior of
SN\,2013dy, SN\,2017cbv, and \abc, the strong similarities between these
events suggest that they may reflect a common physical origin.
